Towards Bringing Together Numerical Methods for Partial Differential Equation and Image Processing

### 2020-03-29 paper/eccv2020submission.tex
Adding a whole bunch of charts

\documentclass{llncs}
\usepackage[title]{appendix}
\usepackage[pagebackref]{hyperref}
\usepackage{subcaption}
\usepackage{graphicx}
\usepackage{comment}
\usepackage{amsmath,amssymb}
\usepackage{color}
\usepackage{float}
\newcommand{\reffig}[1]{\hyperref[#1]{Figure \ref*{#1}}}
\newcommand{\refsec}[1]{\hyperref[#1]{Section \ref*{#1}}}
\newcommand{\reftab}[1]{\hyperref[#1]{Table \ref*{#1}}}
\newcommand{\refapp}[1]{\hyperref[#1]{Appendix \ref*{#1}}}
\DeclareMathOperator*{\argminA}{arg\,min}

@@

\pagestyle{headings}
\mainmatter
\def\ECCVSubNumber{100}
\title{Towards Bringing Together Numerical Methods for Partial Differential Equation and Image Processing}
\author{Stanislav Arnaudov, Markus Hoffmann}
\titlerunning{Towards Bringing Together Numerical Methods for Partial Differential Equation and Deep Neural Networks}
\authorrunning{Stanislav Arnaudov, Markus Hoffmann}
\institute{Karlsruhe Institute of Technology,\\Kaiserstrasse 12,76131 Karlsruhe, Germany\\ \url{http://www.kit.edu/english/}}

@@

The simulations were performed numerically by solving the differential equation describing the flow --- the Navier-Stokes equation. This was done with a numerical solver library --- \emph{HiFlow} [] ---  that works on the base of the Finite element method. The solver supports parallelization with MPI and OpenMP and we used 12 MPI processes to run each simulation. For all of the simulations, the timestep for of solver was set to $0.035$ seconds. This means a single time step of the simulations corresponds to a $0.035$ seconds of physical time.

@@

\subsection{Training approach and networks details}
We base our generative models almost entirely on [pix2pix]. We use the conditional GAN approach to train a generator network that can perform image-to-image translation. As explained in [], the traditional GAN method uses a random vector $z$ as in input to the generator network $G$ to generate output $y$, $G:z \rightarrow y$. Conditional GANs also feed an input image $x$ to the generator, $G: x,z \rightarrow y$. Pix2Pix suggests that in certain cases the usage of $z$ can be usefully but we decided not to include for our generator as we want a deterministic network.

We adopt the objective for the discriminator network as we modify it slightly by leaving out the random vector $z$.
\begin{equation}
\mathcal{L}_{cGAN} (G, D) = (\mathbb{E}[\log D (x,y)] + \mathbb{E}[\log D (x, G(x))])/2
\end{equation}
where $x$ is the input image and $y$ is the target image. In contrast to unconditional GANs, both the generator and the discriminator network have access to the input image. The objective is divided by two to slow down the training of the discriminator relative to the generator as suggested by [].

The objective for the generator network is comprised of two parts --- the value of the discriminator as well as a L1 distance loss between the target and the predicted images. According to [] the L1 loss promotes less blurring and captures the low frequency details of the images. The L1 loss is given by:
\begin{equation}
\mathcal{L}_{L1} (G) = \mathbb{E}[\left\lVert y - G(x)\right\rVert_1]
\end{equation}
The final object for the generator is thus:
\begin{equation}
G^* = \argminA_G \max_D \mathcal{L}_{cGAN}  + \lambda \mathcal{L}_{L1} (G)
\end{equation}
For all of the models we used $\lambda = 100$.

\noindent\textbf{Networks Architectures:} For our generator we use the U-Net variant proposed in pix2pix. It is a standard encoder-decoder model that skip connections between parts of the encoder and the decoder. We also experimented with ResNet based generator but the results were not satisfactory. The network uses blocks of layers of the from convolution-normalization-ReLu. The encoder-decoder first downsamples the input till a bottleneck layer is reached and what follows is an upsampling to the original size of the input image.

@@

For the training procedure, we follow the standard approach proposed by []. With each mini-batch, we first optimize the discriminator and then the generator with the discussed objectives. We use Stochastic Gradient Descent [] with the Adam optimizer with a learning rate of $0.0002$ and momentum parameters $\beta_1=0.9$ and $\beta_2=0.999$.

The used batch size for the constant and inflow speed models was set to 3 and for the viscosity-density models to 4. Those are relatively small numbers for batch sizes but [] suggests that the UNet architecture benefits from small batches in image-to-image translation problems. Our experiments confirmed this.

@@

[Fig] shows the results of the different model types with respect to PSNR.\@ Examples of the predicted images are given in [fig]. We can clearly see that the predicted images are almost identical to the real ones. In all cases, we measured a correlation of about $0.9998$ so we will not give it explicitly here. We think that the images themselves illustrate the high correlation.

[Table] summarizes the results of the evaluation with the average and maximal difference metrics.

@@

  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/single_constant_x}

  \end{subfigure}  
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/single_constant_y}

@@

  \caption{Each image haw tow halves and the upper one shows the ground truth simulation frame. The lower half is the prediction of the model. In the left column, the images are from the x-velocity fie4ld of the simulation and in the right --- form the y-velocity field.}
\end{figure}

\begin{table}
  \begin{center}
    \begin{tabular}{lcc|cc}
      \hline
      \hline
      Model type & \multicolumn{2}{c|}{\underline{Max difference}}  &  \multicolumn{2}{c}{\underline{Average difference}}\\
      \multicolumn{3}{c|}{}&&\\
                 & With pr. & without pr.  &  With pr.& without pr.\\
      \hline
      Constant          &  14.6(3.89)&44.0(3.1e+02)&0.599(0.0818)&0.721(0.294)  \\

@@

      Viscosity-density &  57.0(1.8e+03)&14.7(7.04)&13.8(1.8e+02)&0.374(0.00824) \\
    \end{tabular}
  \end{center}
  \caption{\label{tab:table-name} Your caption.}
\end{table}

@@

The constant model is evaluated with a single simulation --- the one it was trained with. [fig] presents the results of the recursive application. During the evaluation, we noticed that the models can more accurately predict a simulation when the starting frame is further in the simulation. That is, its index relative to the first frame is higher. The plots for the constant model in [fig] illustrate this.

For the inflow speed model, we present the results from three different simulations in [fig]. We can see the same trend that predicting frames with smaller index is harder. We, however, like to note that the networks appear to be handling the parameters properly and adjust the predicted frames accordingly. [fig] presents the results of the viscosity-density model.

@@

We again do not explicitly give the correlation between the predicted and real frames. In all cases, the correlation stayed very close to $1$ and varied in the range $[0.98, 0.99998]$. The CFD-metrics are summarized in [table].

\begin{table}
  \begin{center}
    \begin{tabular}{lcc|cc}
      \hline
      \hline
      Model type & \multicolumn{2}{c|}{\underline{Max difference}}  &  \multicolumn{2}{c}{\underline{Average difference}}\\
      \multicolumn{3}{c|}{}&&\\
                 & With pr. & without pr.  &  With pr.& without pr.\\
      \hline
      Constant          &  14.6(3.89)&44.0(3.1e+02)&0.599(0.0818)&0.721(0.294)  \\

      Viscosity-density &  57.0(1.8e+03)&14.7(7.04)&13.8(1.8e+02)&0.374(0.00824) \\
    \end{tabular}
  \end{center}
  \caption{\label{tab:table-name} Your caption.}
\end{table}

@@

  \caption{These frames are generated by the constant model. The artifacts become apparent after the 10.\@ one.}
\end{figure}

@@

  \caption{Example frames generated by the viscosity-density model. The parameters for this particular simulation are $\nu=4.17647, \rho = 0.001, g=1.4$}
\end{figure}

@@

\section{Networks architectures and hyperparameters}
\label{app1}
\emph{Here we give the exact details on the architectures of the used networks.} The code for all of the models and experiments can be found at \url{https://github.com/palikar/flow_predict}.

The architectures are almost entirely based on~\cite{pix2pix} and [44]. Both the generator and the discriminator networks are comprised of two types of blocks of layers. We will denote a Convolution-BatchNorm-ReLU block with $k$ filters with $Ck$ and Convolution-BatchNorm-Dropout-ReLU with $k$ filters $CDk$. The dropout rate for each dropout layer is 50\%. All convolutions use $4\times4$ filters applied with a stride of $2$.

@@

% LocalWords:  Convolutional timestep timesteps grayscale cGANs
% LocalWords:  preprocessing postprocessing

>>>>

\documentclass{llncs}
\usepackage[title]{appendix}
\usepackage[pagebackref]{hyperref}
\usepackage{subcaption}
\usepackage{graphicx}
\usepackage{comment}
\usepackage{amsmath,amssymb}
\usepackage{color}
\usepackage{float}
\newcommand{\reffig}[1]{\hyperref[#1]{Figure~\ref*{#1}}}
\newcommand{\refsec}[1]{\hyperref[#1]{Section~\ref*{#1}}}
\newcommand{\reftab}[1]{\hyperref[#1]{Table~\ref*{#1}}}
\newcommand{\refapp}[1]{\hyperref[#1]{Appendix~\ref*{#1}}}
\newcommand{\thickhline}{%
    \noalign {\ifnum 0=`}\fi \hrule height 1pt
    \futurelet \reserved@a \@xhline
}
\DeclareMathOperator*{\argminA}{arg\,min}

@@

\pagestyle{headings}
\mainmatter\def\ECCVSubNumber{100}
\title{Towards Bringing Together Numerical Methods for Partial Differential Equation and Image Processing}
\author{Stanislav Arnaudov, Markus Hoffmann}
\titlerunning{Towards Bringing Together Numerical Methods for Partial Differential Equation and Deep Neural Networks}
\authorrunning{Stanislav Arnaudov, Markus Hoffmann}
\institute{Karlsruhe Institute of Technology,\\Kaiserstrasse 12,76131 Karlsruhe, Germany\\ \url{http://www.kit.edu/english/}}

@@

The simulations were performed numerically by solving the differential equation describing the flow --- the Navier-Stokes equation. This was done with a numerical solver library --- \emph{HiFlow}~\cite{gawlok2017} ---  that works on the base of the Finite element method. The solver supports parallelization with MPI and OpenMP and we used 12 MPI processes to run each simulation. For all of the simulations, the timestep for of solver was set to $0.035$ seconds. This means a single time step of the simulations corresponds to a $0.035$ seconds of physical time.

@@

\subsection{Training approach and networks details}
We base our generative models almost entirely on~\cite{pix2pix}. We use the conditional GAN approach to train a generator network that can perform image-to-image translation. As explained in~\cite{pix2pix}, the traditional GAN method uses a random vector $z$ as in input to the generator network $G$ to generate output $y$, $G:z \rightarrow y$. Conditional GANs also feed an input image $x$ to the generator, $G: x,z \rightarrow y$.~\cite{pix2pix} and~\cite{wang2016} suggest that in certain cases the usage of $z$ can be usefully but we decided not to include for our generator as we want a deterministic network.

We adopt the objective for the discriminator network as we modify it slightly by leaving out the random vector $z$.
\begin{equation}
\mathcal{L}_{cGAN} (G, D) = (\mathbb{E}[\log D (x,y)] + \mathbb{E}[\log D (x, G(x))])/2
\end{equation}
where $x$ is the input image and $y$ is the target image. In contrast to unconditional GANs, both the generator and the discriminator network have access to the input image. The objective is divided by two to slow down the training of the discriminator relative to the generator as suggested by~\cite{pix2pix}.

The objective for the generator network is comprised of two parts --- the value of the discriminator as well as a L1 distance loss between the target and the predicted images. According to~\cite{pix2pix} the L1 loss promotes less blurring and captures the low frequency details of the images. The L1 loss is given by:
\begin{equation}
\mathcal{L}_{L1} (G) = \mathbb{E}[\left\lVert y - G(x)\right\rVert_1]
\end{equation}
The final object for the generator is thus:
\begin{equation}
G^* = \argminA_G \max_D \mathcal{L}_{cGAN}  + \lambda \mathcal{L}_{L1} (G)
\end{equation}
For all of the models we used $\lambda = 100$.

\noindent\textbf{Networks Architectures:} For our generator we use the U-Net~\cite{ronneberger2015} variant proposed in~\cite{pix2pix}. It is a standard encoder-decoder~\cite{hinton2006} model that skip connections between parts of the encoder and the decoder. We also experimented with ResNet based generator but the results were not satisfactory. The network uses blocks of layers of the from convolution-normalization-ReLu. The encoder-decoder first downsamples the input till a bottleneck layer is reached and what follows is an upsampling to the original size of the input image.

@@

For the training procedure, we follow the standard approach proposed by~\cite{goodfellow2014}. With each mini-batch, we first optimize the discriminator and then the generator with the discussed objectives. We use Stochastic Gradient Descent~\cite{kiefer1952} with the Adam optimizer~\cite{kingma2014} with a learning rate of $0.0002$ and momentum parameters $\beta_1=0.9$ and $\beta_2=0.999$.
The used batch size for the constant and inflow speed models was set to 3 and for the viscosity-density models to 4. Those are relatively small numbers for batch sizes but~\cite{pix2pix} suggests that the U-Net architecture benefits from small batches in image-to-image translation problems. Our experiments confirmed this.

@@

\reffig{fig:single_psnr} shows the results of the different model types with respect to PSNR.\@ Examples of the predicted images are given in \reffig{fig:single_images}. We can clearly see that the predicted images are almost identical to the real ones. In all cases, we measured a correlation of about $0.9998$ so we will not give it explicitly here. We think that the images themselves illustrate the high correlation.

\reftab{tab:single} summarizes the results of the evaluation with the average and maximal difference metrics.

@@

  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/single_constant_x}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/single_constant_y}

@@

  \caption{Each image haw tow halves and the upper one shows the ground truth simulation frame. The lower half is the prediction of the model. In the left column, the images are from the x-velocity fie4ld of the simulation and in the right --- form the y-velocity field.}\label{fig:single_images}
\end{figure}

\begin{figure}
  \begin{subfigure}{.3\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/charts/single_const_psnr}
  \end{subfigure}
  \begin{subfigure}{.3\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/charts/single_fluid_psnr}
  \end{subfigure}
  \begin{subfigure}{.3\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/charts/single_speed_psnr}
  \end{subfigure}

  \caption{}\label{fig:single_psnr}
\end{figure}

\begin{table}
  \begin{center}
    \begin{tabular}{lcc|cc}

      \textbf{Model type} & \multicolumn{2}{c|}{\textbf{Max difference}}  &  \multicolumn{2}{c}{\textbf{Average difference}}\\
      \hline 
      \multicolumn{3}{c|}{}&&\\
                 & \underline{With pr.} & \underline{without pr.}  &  \underline{With pr}.& \underline{without pr.}\\

      Constant          &  14.6(3.89)&44.0(3.1e+02)&0.599(0.0818)&0.721(0.294)  \\

@@

      Viscosity-density &  57.0(1.8e+03)&14.7(7.04)&13.8(1.8e+02)&0.374(0.00824) \\
    \end{tabular}
  \end{center}
  \caption{Your caption.}\label{tab:single}
\end{table}

@@

The constant model is evaluated with a single simulation --- the one it was trained with. \reffig{fig:rec_const_psnr} presents the results of the recursive application. During the evaluation, we noticed that the models can more accurately predict a simulation when the starting frame is further in the simulation. That is, its index relative to the first frame is higher. The plots for the constant model in \reffig{fig:rec_const_psnr} illustrate this.

For the inflow speed model, we present the results from two different simulations in \reffig{fig:rec_speed_psnr}. We can see the same trend that predicting frames with smaller index is harder. We, however, like to note that the networks appear to be handling the parameters properly and adjust the predicted frames accordingly.\reffig{fig:rec_fluid_psnr} presents two simulations of the viscosity-density model.

@@

We again do not explicitly give the correlation between the predicted and real frames. In all cases, the correlation stayed very close to $1$ and varied in the range $[0.98, 0.99998]$. The CFD-metrics are summarized in [].

@@

  \caption{These frames are generated by the constant model. The artifacts become apparent after the 10.\@ one.}\label{fig:const_sim}
\end{figure}

@@

  \caption{Example frames generated by the viscosity-density model. The parameters for this particular simulation are $\nu=4.17647, \rho = 0.001, g=1.4$}\label{fig:fluid_sim}
\end{figure}

\begin{table}
  \begin{center}
    \begin{tabular}{l|cc|cc}
      \textbf{Frame} & \multicolumn{2}{c|} {\shortstack{\textbf{Simulation}\\$\nu=4.17647, \rho = 0.001, g=1.4$}}  &  \multicolumn{2}{c}{\shortstack{\textbf{Simulation}\\$\nu=4.17647, \rho = 0.001, g=1.4$}}\\
      \hline 
                     &\multicolumn{2}{c|}{} &&\\
                     & {Max difference} & {Average difference}  &  {Max difference} & {Average difference}\\
      \emph{Frame 1}       &35    &35     &35    &35\\
      \emph{Frame 5}       &35    &35     &35    &35\\
      \emph{Frame 10}      &35    &35     &35    &35\\
      \emph{Frame 15}      &35    &35     &35    &35\\
      \emph{Frame 20}      &35    &35     &35    &35\\
      \emph{Frame 25}      &35    &35     &35    &35\\
      \emph{Frame 30}      &35    &35     &35    &35\\
      \emph{Frame 35}      &35    &35     &35    &35\\
      \emph{Frame 40}      &35    &35     &35    &35\\

    \end{tabular}
  \end{center}
  \caption{Your caption.}\label{tab:recursive}
\end{table}

\begin{figure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/charts/constant_i10}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/charts/constant_i190}
  \end{subfigure}  
  \caption{}\label{fig:rec_const_psnr}
\end{figure}

\begin{figure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/charts/inflow_i10}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/charts/inflow_i1436}
  \end{subfigure}

  \begin{center}
    Simulation
  \end{center}

  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/charts/inflow_39_s0}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/charts/inflow_39_s120}
  \end{subfigure}

  \begin{center}
    Simulation
  \end{center}

  \caption{}\label{fig:rec_speed_psnr}
\end{figure}

\begin{figure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/charts/fluid_25_s0}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/charts/fluid_25_s120}
  \end{subfigure}

  \begin{center}
    Simulation
  \end{center}

  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/charts/fluid_100_s0}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/charts/fluid_100_s120}
  \end{subfigure}

  \begin{center}
    Simulation
  \end{center}

  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/charts/fluid_300_s0}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{imgs/charts/fluid_300_s120}
  \end{subfigure}

  \begin{center}
    Simulation
  \end{center}

  \caption{}\label{fig:rec_fluid_psnr}
\end{figure}

@@

\section{Networks architectures and hyperparameters}\label{app1}
\emph{Here we give the exact details on the architectures of the used networks.} The code for all of the models and experiments can be found at \url{https://github.com/palikar/flow_predict}.

The architectures are almost entirely based on~\cite{pix2pix} and~\cite{radford2015}. Both the generator and the discriminator networks are comprised of two types of blocks of layers. We will denote a Convolution-BatchNorm-ReLU block with $k$ filters with $Ck$ and Convolution-BatchNorm-Dropout-ReLU with $k$ filters $CDk$. The dropout rate for each dropout layer is 50\%. All convolutions use $4\times4$ filters applied with a stride of $2$.

@@

% LocalWords:  Convolutional timestep timesteps grayscale cGANs GAN
% LocalWords:  preprocessing postprocessing hyperparameters HiFlow
